Multitarget Development System (MDS) for PicoBlaze is a graphical integrated development environment (IDE) for
Xilinx's PicoBlaze soft-core processors and their compatible clones. MDS is intended to be used mainly by professional
developers and by education institutions, for hobbyists there is also noncommercial version available under a different
license.

MDS provides all the necessary functionality to develop software part of a PicoBlaze application, including source code
editor, assembler, disassembler, and simulator. Besides that there is also a number of features and functions to make
your work easier, the sole purpose of MDS is to save your time and enable development of more complex applications. User
is our main concern. Having the user as the center of attention, we developed this IDE to be intuitive, fast and rich in
features. We prefer to use state of the art technologies to keep up with the pace of technological development in this
area of industry, and MDS benefits from such approach. We believe you will feel comfortable while working with this tool.

\section{Features}
    \begin{itemize}
        \item Text editor optimized for writing source code.
        \item Project manager for creating and maintaining your projects
        \item Fast simulator for all available versions of Picoblaze.
        \item Macro-assembler supporting wide range of output file formats.
        \item Disassembler.
        \item Tool called Assembler Translator for compatibility with your current tools.
        \item Command line tools, and a number of graphical utilities.
    \end{itemize}

\section{Requirements}
    This software is compiled for x86 processors in both 32b and 64b variant with SSE instruction set that means you
    need a processor equipped with SEE feature (most processors manufactured after the year 2003 are). For GNU/Linux,
    MDS requires GLIBC version >= 2.14.

    Targeted operating systems are:
    \begin{itemize}
        \item Microsoft Windows 7 (32b and 64b),
        \item Microsoft Windows XP (32b),
        \item Ubuntu 13.10 and higher (32b and 64b),
        \item openSUSE 13.1 and higher (32b and 64b),
        \item Fedora 20 and higher (32b and 64b),
        \item CentOS 7 (64b).
    \end{itemize}
