\definecolor{psinstruction_bg}{rgb}{0.6, 0.9, 1.0}
\newcommand{\psinstruction}[1]{~\\[7pt]\addcontentsline{toc}{subsection}{#1}\colorbox{psinstruction_bg}{\parbox{\dimexpr\textwidth-2\fboxsep}{\color{black}\textbf{#1}}}\bigskip}

MDS assembler supports a number of pseudo instructions which can improve understandability of your source code. Assembler will replace these instructions with one or more PicoBlaze instructions to achieve the purpose of that pseudo-instruction.

\psinstruction{NOP}
    \index{NOP}
    No operation. This instruction does not do anything, it just consumes processor time.

    \paragraph{Syntax}
        ~\\
        \verb'    NOP'

    \paragraph{Equivalent}
        ~\\
        \verb'    LOAD  s0, s0'

    \paragraph{Examples}
        ~\\
        \verb'    NOP'\\
        \verb'    NOP'\\
        \verb'    NOP'

\psinstruction{INC}
    \index{INC}
    Increments the given register value by one.

    \paragraph{Syntax}
        ~\\
        \verb'    INC   sX'

    \paragraph{Equivalent}
        ~\\
        \verb'    ADD   sX, #1'

    \paragraph{Examples}
        ~\\
        \verb'    INC   s0'

\clearpage
\psinstruction{DEC}
    \index{DEC}
    Decrements the given register value by one.

    \paragraph{Syntax}
        ~\\
        \verb'    DEC   sX'

    \paragraph{Equivalent}
        ~\\
        \verb'    SUB   sX, #1'

    \paragraph{Examples}
        ~\\
        \verb'    DEC   s0'

\psinstruction{SETR}
    \index{SETR}
    Sets the given register to binary ones.

    \paragraph{Syntax}
        ~\\
        \verb'    SETR  sX'

    \paragraph{Equivalent}
        ~\\
        \verb'    OR    sX, #0xFF'

    \paragraph{Examples}
        ~\\
        \verb'    SETR  s0'

\psinstruction{CLRR}
    \index{CLRR}
    Clears the given register (set it to 0).

    \paragraph{Syntax}
        ~\\
        \verb'    CLRR  sX'

    \paragraph{Equivalent}
        ~\\
        \verb'    AND   sX, #0'

    \paragraph{Examples}
        ~\\
        \verb'    CLRR  s0'

\clearpage
\psinstruction{CPL}
    \index{CPL}
    Performs ones' complement with the given register.

    \paragraph{Syntax}
        ~\\
        \verb'    CPL  sX'

    \paragraph{Equivalent}
        ~\\
        \verb'    XOR   sX, #0xFF'

    \paragraph{Examples}
        ~\\
        \verb'    CPL   s0'

\psinstruction{CPL2}
    \index{CPL2}
    Performs two's complement with the given register.

    \paragraph{Syntax}
        ~\\
        \verb'    CPL2  sX'

    \paragraph{Equivalent}
        ~\\
        \verb'    XOR   sX, #0xFF'\\
        \verb'    ADD   sX, #1'

    \paragraph{Examples}
        ~\\
        \verb'    CPL2  s0'

\psinstruction{SETB}
    \index{SETB}
    Sets one bit in the given register.

    \paragraph{Syntax}
        ~\\
        \verb'    SETB  sX, bit ; bit belongs to interval [0,7]'

    \paragraph{Equivalent}
        ~\\
        \verb'    OR    sX, # 1 << bit'

    \paragraph{Examples}
        ~\\
        \verb'    SETB  S0, 3'

\clearpage
\psinstruction{CLRB}
    \index{CLRB}
    Clears one bit in the given register.

    \paragraph{Syntax}
        ~\\
        \verb'    CLRB  sX, bit ; bit belongs to interval [0,7]'

    \paragraph{Equivalent}
        ~\\
        \verb'    AND   sX, # ( 0xFF ^ ( 1 << bit ) )'

    \paragraph{Examples}
        ~\\
        \verb'    CLRB  S0, 7'

\psinstruction{NOTB}
    \index{NOTB}
    Negates one bit in the given register.

    \paragraph{Syntax}
        ~\\
        \verb'    NOTB  sX, bit ; bit belongs to interval [0,7]'

    \paragraph{Equivalent}
        ~\\
        \verb'    XOR   sX, # ( ~( 1 << bit ) )'

    \paragraph{Examples}
        ~\\
        \verb'    NOTB  S0, 7'

\psinstruction{DJNZ}
    \index{DJNZ}
    Decrements the given register and jumps at the given label until the register contains zero.

    \paragraph{Syntax}
        ~\\
        \verb'    DJNZ  sX, label'

    \paragraph{Equivalent}
        ~\\
        \verb'    SUB   sX, #1'\\
        \verb'    JUMP  NZ, label'

    \paragraph{Examples}
        ~\\
        \verb'    loop:             '\\
        \verb'        DJNZ  s0, loop'

\clearpage
\psinstruction{IJNZ}
    \index{IJNZ}
    Increments the given register and jumps at the given label until the register contains zero.

    \paragraph{Syntax}
        ~\\
        \verb'    IJNZ sX, label'

    \paragraph{Equivalent}
        ~\\
        \verb'    ADD   sX, #1'\\
        \verb'    JUMP  NZ, label'

    \paragraph{Examples}
        ~\\
        \verb'    loop:             '\\
        \verb'        IJNZ  s0, loop'
